Simulation of Nano Particles in a Laser Trap

### 2017-02-17 doc/draft3.tex
updates

\begin{document}
\title{Simulation of Nano Particles in a Laser Trap}
\author{Mathias H\"old, BSc.}
\date{2016}
\maketitle
\thispagestyle{empty}
\newpage
\tableofcontents
\newpage
\section{Introduction}
%something about computer science in general, optical tweezers and the need to study the
%systems on computers yada yada
In the year 1953, Metropolis et al.~\cite{Metropolis1953} used the MANIAC (Mathematical Analyzer Numerical Integrator And Computer) 
at the Los Alamos National Laboratories in the United States to
simulate particles in a liquid state to calculate the equation of state. This was the first publication in which scientists used a computer to
calculate numerical results based on a physical theory in a process that is to this day called \textit{Metropolis Monte Carlo}. To say that a lot has
changed in the field of computer simulation (or computers in general) would be a vast understatement.\\
Not only the processing power of computers has incresed exponentially (Moore's law), which allows the simulation of a greater number of particles for
a longer period of time, but also new techniques have been developed, such as molecular dynamics. These techniques are not only used by physicists,
but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. This thesis is an
attempt to apply methods of simulation to a field with a lot of recent research happening in it: optical trapping.
Optical trapping is a technique that has been used and developed for over 40 years~\cite{Marago2013}. The idea for this goes back to the beginning of
the 20th century, where Lebedev~\cite{Lebedew1901}, Nichols and Hull~\cite{Nichols1901} demonstrated the existence of a force that light exerts on 
matter. Due to the technical limitations at that time, the idea was discarded until 1970, where Ashkin~\cite{Ashkin2000} used lasers to manipulate the
motion of neutral atoms and micrometer sized objects and thus brought the idea of optical trapping back to life.\\
With this advancement it was possible to develop methods for trapping and manipulating objects of different sizes ranging from the micrometer scale
down to the subnano scale. At first, this technique was used to either trap objects in the micrometer range (like cells) or in the subnano scale,
where lasers were used for the cooling of atoms. The range between those two, with objects at the scale of one to a hundred nanometer, was not
explored as much because neither downscaling the methods for micrometer sized objects nor upscaling the methods for atoms seemed feasible. In recent
years however, new developments in the field opened the possibilities to work with a variety of objects in the desired nanometer scale, such as metal
nanoparticles, nano tubes, quantum dots and much more. This non-invasive method of trapping objects allows very precise measurements of various
properties and also allows the cooling of nanoparticles towards their quantum mechanical ground state.\\
The small size of the objects used in the optical tweezer experiments make them ideal candidates for computer simulations. In this thesis, I will
model an optical tweezer setup, based on a real experiment, to investigate different temperatures in a non-equilibrium state.\\
The first chapter will be the introduction of the experiment and the introduction of the problem statement. The second chapter contains basic
information on the techniques used to model the experiment and simulate the dynamic properties. In the third chapter, I will present my findings which
are followed by the conclusions and further remarks.

@@

In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$kg is trapped in a laser beam within a

@@

and the different temperatures are depicted in Fig.~\ref{fig:levitation}.\\

@@

            of the actual experimental setup. In \textbf{c} the 4 different temperatures that are present during the experiment are depicted and
        \textbf{d} shows the position of the particle in the trap over time. The image was taken from~\cite{MillenJ.2014}}

@@

equilibrium, which is in contrast to the assumption in many optical tweezer experiments. Since the gas particles only interact with the nano particle
in the trap and not wich each other, this situation creates two heat baths of different temperatures with the nano particle acting as a mediator

@@

    different temperatures. Image modified from \cite{Kroy2014}}

@@

\newpage
\section{Simulation}
%simulation techniques and general concept of the simulation itself, used methods and so on 
The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
that are widely used, one of which being molecular dynamics (MD) simulations.\\ 
The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
into individual pieced that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on an FCC
lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, as it acts as a trapping device as well
as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, that acts as a thermostat in the simulation, that will be
modeled as a box surrounding the nano particle. This box is filled with gas particles that interact with the nano particle via a soft-sphere potential
and do not interact with one another.\\
The following section will give a brief overview of the concepts used to simulate every part of the experiment.

@@

Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$m and $\varepsilon = 1.65 \times 10^{-21}$J.\\

@@

With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell 

@@

            \caption{Schematic figure of a face centered cubic (FCC) lattice}

@@

            \caption{Setup for successively building a FCC lattice}

@@

where $L$ is the side length of the cube that is the whole system and it is calculated via the density of the system

@@

expansion of the coordinates mathbftor \eqref{eq:verletnew} and instead of Taylor expanding $\mathbf{r}_i(t-\Delta t)$, we write it as 

@@

notation of Fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 

@@

where $\varepsilon$ is the interaction strength and $\sigma$ is the interaction range and $r$ is the distance between the gas particle and the

@@

        If the length of the cell $L$ is chosen to be equal to the cut-off radius $r_c$, this insertion should only be carried out with a probablity of

@@

        \caption{Instantaneous and mean temperature of the outgoing gas particles of the surrounding gas}

@@

        \caption{Instantaneous and mean temperature of the center of mass of the nano particle in the laser trap}

@@

        \caption{Instantaneous and mean internal temperature of the nano particle in the laser trap}

@@

        \caption{A histogram of the x-component of the center of mass velocities with the expected gaussian distribution}

@@

The first thing that is noticable, is the increase of the temperature of the center of mass, the internal temperature and the outgoing gas temperature

@@

\newpage
\section{Conclusion and Further Remarks}
% what does this all tell us? do we need to study more of this? 
In this thesis I have modeled an optical tweezer experiment with methods from the field of computational physics. The experiment was broken down into
different sections, that have been modeled separately. With the use of molecular dynamics I simulated the behaviour of a nano particle in a laser trap
and investigated the effects of the laser power and temperature of the surrounding gas on the center of mass movement of the nano particle.\\
The results show, that the center of mass temperature (and thus its movement) does indeed depend on the energy of the laser, as well as on the
pressure and temperature of the surrounding gas.\\
The development of the center of mass temperature shows a curve that seems to have a convex form, whereas the internal temperature shows a more
concave behaviour. So while the internal temperature initially grows faster, it could be possible for the center of mass temperature to pass the
internal temperature.\\
There are several ways to explore this subject further. The first idea would be to decrease the pressure even more and investiate the case of very low
energies, since this is not easily feasible in an experimental setup. Another idea would be to increase the number of atoms in the nano particle to
see if the surface of the particle plays a role in the development of the temperatures. Furthermore, the interaction between the gas particles and the
nano particle in the trap could be varied to include an attractive part in the potential. Generally, these simulations could be expanded by running
them on faster computers to investigate long term behaviour of the systems and expand the ranges of the different values like pressure or laser
energy.

>>>>

\begin{document}
\title{Simulation of Nano Particles in a Laser Trap}
\author{Mathias H\"old, BSc.}
\date{2016}
\maketitle
\thispagestyle{empty}
\newpage
\tableofcontents
\newpage
\section{Introduction}
%something about computer science in general, optical tweezers and the need to study the
%systems on computers yada yada
In the year 1953, Metropolis et al.~\cite{Metropolis1953} used the MANIAC (Mathematical Analyzer Numerical Integrator And Computer) 
at the Los Alamos National Laboratories in the United States to
simulate particles in a liquid state to calculate the equation of state. This was the first publication in which scientists used a computer to
calculate numerical results based on a physical theory in a process that is to this day called \textit{Metropolis Monte Carlo}. To say that a lot has
changed in the field of computer simulation (or computers in general) would be a vast understatement.\\
Not only has the processing power of computers increased exponentially (Moore's law), which allows the simulation of a greater number of particles for
a longer period of time, but also new techniques have been developed, such as molecular dynamics. These techniques are not only used by physicists,
but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. This thesis is an
attempt to apply methods of simulation to a field of active research: optical trapping.
Optical trapping is a technique that has been used and developed for over 40 years~\cite{Marago2013}. The idea for this goes back to the beginning of
the 20th century, where Lebedev~\cite{Lebedew1901}, Nichols and Hull~\cite{Nichols1901} demonstrated the existence of a force that light exerts on 
matter. Due to the technical limitations at that time, the idea was discarded until 1970, where Ashkin~\cite{Ashkin2000} used lasers to manipulate the
motion of neutral atoms and micrometer sized objects and thus brought the idea of optical trapping back to life.\\
With this advancement it was possible to develop methods for trapping and manipulating objects of different sizes ranging from the micrometer scale
down to the subnano scale. At first, this technique was used to either trap objects in the micrometer range (like cells) or in the subnano scale,
where lasers were used for the cooling of atoms. The range between those two, with objects at the scale of one to a hundred nanometer, was not
explored as much because neither downscaling the methods for micrometer sized objects nor upscaling the methods for atoms seemed feasible. In recent
years however, new developments in the field opened the possibilities to work with a variety of objects in the desired nanometer scale, such as metal
nanoparticles, nano tubes, quantum dots and much more. This non-invasive method of trapping objects allows very precise measurements of various
properties and also allows the cooling of nanoparticles towards their quantum mechanical ground state.\\
The small size of the objects used in the optical tweezer experiments make them ideal candidates for computer simulations. In this thesis, I will
model an optical tweezer setup, based on a real experiment, to investigate different temperatures in a non-equilibrium state.\\
The first chapter will be the introduction of the experiment and the introduction of the problem statement. The second chapter contains basic
information on the techniques used to model the experiment and simulate the dynamic properties. In the third chapter, I will present my findings which
are followed by the conclusions and further remarks.

@@

In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$ kg is trapped in a laser beam within a

@@

and the different temperatures are shown in Fig.~\ref{fig:levitation}.\\

@@

            of the actual experimental setup. In \textbf{c} the 4 different temperatures that are present during the experiment are illustrated and
        \textbf{d} shows the position of the particle in the trap over time. The image was taken from~\cite{MillenJ.2014}.}

@@

equilibrium, which is in contrast to the assumptions made in many optical tweezer experiments. Since the gas particles only interact with the nano particle
in the trap and not which each other, this situation creates two heat baths of different temperatures with the nano particle acting as a mediator

@@

    different temperatures. Image modified from \cite{Kroy2014}.}

@@

\newpage
\section{Simulation}
%simulation techniques and general concept of the simulation itself, used methods and so on 
The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
that are widely used, one of which being molecular dynamics (MD) simulations.\\ 
The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
into individual pieces that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on a FCC
(face-centered cubic) lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, 
as it acts as a trapping device as well as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, 
that acts as a thermostat in the simulation, which will be modeled as a cubic box surrounding the nano particle. 
This box is filled with gas particles that interact with the nano particle via a soft-sphere potential and do not interact with one another.\\
The following section will give a brief overview of the concepts used to simulate every part of the experiment.

@@

Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$ m and $\varepsilon = 1.65 \times 10^{-21}$ J.\\

@@

With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since one FCC cell 

@@

            \caption{Schematic figure of a face-centered cubic (FCC) lattice.}

@@

            \caption{Setup for successively building a FCC lattice.}

@@

where $L$ is the side length of the cube that represents the nano particle and it is calculated via the density of the system

@@

expansion of the coordinates vector \eqref{eq:verletnew} and instead of Taylor expanding $\mathbf{r}_i(t-\Delta t)$, we write it as 

@@

notation of Fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither the center of mass of the particles the simulation box nor 

@@

where $\varepsilon$ is the interaction strength, $\sigma$ is the interaction range and $r$ is the distance between the gas particle and the

@@

        If the length of the cell $L$ is chosen to be equal to the cut-off radius $r_c$, this insertion should only be carried out with a probability of

@@

        \caption{Instantaneous and mean temperature of the outgoing gas particles of the surrounding gas.}

@@

        \caption{Instantaneous and mean temperature of the center of mass of the nano particle in the laser trap.}

@@

        \caption{Instantaneous and mean internal temperature of the nano particle in the laser trap.}

@@

        \caption{A histogram of the x-component of the center of mass velocities with the expected gaussian distribution.}

@@

The first thing that is noticeable, is the increase of the temperature of the center of mass, the internal temperature and the outgoing gas temperature

@@

\newpage
\section{Conclusion and Further Remarks}
% what does this all tell us? do we need to study more of this? 
In this thesis I have modeled an optical tweezer experiment with methods from the field of computational physics. The experiment was broken down into
different sections, that have been modeled separately. With the use of molecular dynamics I simulated the behaviour of a nano particle in a laser trap
and investigated the effects of the laser power and temperature of the surrounding gas on the center of mass movement of the nano particle.\\
The results show, that the center of mass temperature (and thus its movement) does indeed depend on the energy of the laser, as well as on the
pressure and temperature of the surrounding gas.\\
The development of the center of mass temperature shows a curve that seems to have a convex form, whereas the internal temperature shows a more
concave behaviour. So while the internal temperature initially grows faster, it could be possible for the center of mass temperature to pass the
internal temperature.\\
There are several ways to explore this subject further. The first idea would be to decrease the pressure even more and investigate the case of very low
energies, since this is not easily feasible in an experimental setup. Another idea would be to increase the number of atoms in the nano particle to
see if the surface of the particle plays a role in the development of the temperatures. Furthermore, the interaction between the gas particles and the
nano particle in the trap could be varied to include an attractive part in the potential. Generally, these simulations could be expanded by running
them on faster computers to investigate long term behaviour of the systems and expand the ranges of the different values like pressure or laser
energy.
